
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(ggplot2)
library(rjson)
library(tidyverse)
library(dplyr)

data2000 <- read.csv("./period_2000_rearranged.csv",header = TRUE) %>% select(-TIME_PERIOD,-PTP_RT,-PTP_TIME)
      data2000$TIME_SP <- as.numeric(lubridate::hm(data2000$TIME_SP) %>% lubridate::seconds())
      
data2010 <- read.csv("./period_2010_rearranged.csv",header = TRUE) %>% select(-TIME_PERIOD,-PTP_RT,-PTP_TIME)
      data2010$TIME_SP <- as.numeric(lubridate::hm(data2010$TIME_SP) %>% lubridate::seconds())
      
dataLabeledByActivityType <- rbind(data2000,data2010) %>% group_by(sex,age,acl00,geo) %>% summarise(TIME_SP = mean(TIME_SP))
dataLabeledByActivityType <- dataLabeledByActivityType[dataLabeledByActivityType$sex != "T",]


dataLabeledByActivityType[dataLabeledByActivityType == 0] = NA

labelsAndActivities <- fromJSON(file = "./CodeLists/ACL00.json") %>% flatten() %>% bind_rows()

dataLabeledByActivityType <- full_join(dataLabeledByActivityType,labelsAndActivities)
dataLabeledByActivityType <- dataLabeledByActivityType[!is.na(dataLabeledByActivityType$label) & !is.na(dataLabeledByActivityType$TIME_SP),]
dataLabeledByActivityType$label <- mapply(FUN = str_wrap,dataLabeledByActivityType$label,width = 15)

  
```

## The Dataset
The data used in this small project is the result of the Time Use Survey(TUS) which had in mind the measurement of the amount of time people spent on various activities throughout the day. data is filled by individuals who keep a diary that they can fill, so this data is self reported and should be treated as such.  
The dataset contains data from two waves of those surveys, part of the Harmonized European Time Use Survey(HETUS). First wave containing 15 countries with the interviews taking place between 1998 and 2006, referred as the 2000 period from now on, and the second wave taking place between 2008 and 2015 in 18 countries, from now on referred as the 2010 period.
What we were interested from this dataset was to see specifically the kind of main activities, found as the Activity Coding List(ACL) in the dataset, people usually participated in and about how much time individuals who participated in certain activities spent.  
The diaries that the participant were supposed to complete had every 24h day that they were supposed to fill in split into 144 slots of 10 minutes each of those slots were marked with one of 108 categories of activities, of which they ended up reduced as 51 different categories.   


## A Few Observations

Here we have a graph that shows the mean time spent by Male and Female counterparts on different types of activities:

```{r all_age_split_on_activity, echo=FALSE, fig.height=15,error=FALSE,warning=FALSE}
ggplot(data = dataLabeledByActivityType,aes(x = sex, y = TIME_SP, fill=sex)) + geom_bar(stat = "summary",position = "dodge") + 
  ylab("Time in Seconds")   +  
  facet_grid(label ~ sex,scales = "free") + xlab("average of all ages") +  
  theme(strip.text = element_text(size = 15),
                  axis.title.x = element_text(size = 15),
                  axis.title.y = element_text(size = 15),
                  axis.text.y = element_text(size = 15),
                  axis.text.x = element_text(size = 15),
                  panel.spacing.y = unit(1,"cm",data = NULL)
                  ,legend.position="none")
```  

As we can see from the above graph, in general women spend more time on activities related to family care, as childcare, and on work that entails maintaining a household, while males usually spend more time at work or work related activities  

It might be useful to have a closer look at how people are using their time depending on age groups, as we have observed that differences can be quite big depending on the age of the volunteers  

```{r 20_t0_24_split_on_activity, echo=FALSE, fig.height=15,error=FALSE,warning=FALSE}
ggplot(data = dataLabeledByActivityType[dataLabeledByActivityType$age == "Y20-24",],aes(x = sex, y = TIME_SP, fill=sex)) + geom_bar(stat = "summary",position = "dodge") + 
  ylab("Time in Seconds")   +  xlab("20 to 24") +  
  facet_grid(label ~ sex,scales = "free") +
  theme(strip.text = element_text(size = 15),
                  axis.title.x = element_text(size = 15),
                  axis.title.y = element_text(size = 15),
                  axis.text.y = element_text(size = 15),
                  axis.text.x = element_text(size = 15),
                  panel.spacing.y = unit(1,"cm",data = NULL)
                  ,legend.position="none")
```   
  
Maybe not unexpectedly younger people do seem to spend way more time on activities related to education and less on those related to family time and house keeping
  
```{r 25_t0_44_split_on_activity, echo=FALSE, fig.height=15,error=FALSE,warning=FALSE}
ggplot(data = dataLabeledByActivityType[dataLabeledByActivityType$age == "Y25-44",],aes(x = sex, y = TIME_SP, fill=sex)) + geom_bar(stat = "summary",position = "dodge") + 
  ylab("Time in Seconds")   +  xlab("25 to 44") +  
  facet_grid(label ~ sex,scales = "free") +
  theme(strip.text = element_text(size = 15),
                  axis.title.x = element_text(size = 15),
                  axis.title.y = element_text(size = 15),
                  axis.text.y = element_text(size = 15),
                  axis.text.x = element_text(size = 15),
                  panel.spacing.y = unit(1,"cm",data = NULL)
                  ,legend.position="none")
```  
  
As age progresses males keep increasing the time spent with family having an increase of about 30% while women who where already spending almost twice as much time on those activities than their male counterpart further increase the difference by almost doubling their previous average time spent  
  

```{r 45_t0_64_split_on_activity, echo=FALSE, fig.height=15,error=FALSE,warning=FALSE}
ggplot(data = dataLabeledByActivityType[dataLabeledByActivityType$age == "Y45-64",],aes(x = sex, y = TIME_SP, fill=sex)) + geom_bar(stat = "summary",position = "dodge") + 
  ylab("Time in Seconds")   + xlab("45 to 64") +  
  facet_grid(label ~ sex,scales = "free") +
  theme(strip.text = element_text(size = 15),
                  axis.title.x = element_text(size = 15),
                  axis.title.y = element_text(size = 15),
                  axis.text.y = element_text(size = 15),
                  axis.text.x = element_text(size = 15),
                  panel.spacing.y = unit(1,"cm",data = NULL)
                  ,legend.position="none")
```  
  
From the few above graphs it can be seen that usually activities related to pleasure and time for oneself stay relatively stable, activities related to education sharply decrease after the age of 24 for both males and females. It seems that women end up spending about double the time in activities related to housekeeping and spending time with their family while men consistently spend more time on activities related to employment

Since our chosen dataset's main classification system used  is the "Activity coding list" (ACL), which assigns a code to the activities (primary and secondary) that the person is doing during the day, we decided to present a graph on the primary activities. 
The graph shows mean time spent by Male and Female counterparts on the primary types of activities:

```{r maligraph2, echo=FALSE, fig.height=15,error=FALSE,warning=FALSE}
#male activity time spent
library(gridExtra)
period_2000_rearranged <- read.csv("./period_2000_rearranged.csv")
Graph1 <- data.frame(period_2000_rearranged)
data_AC0 = subset(Graph1,sex == "M" & acl00 =="AC0" & age =="TOTAL")
time1 <- data_AC0$TIME_SP
my_list_AC0 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC0 <- mean(my_list_AC0)

data_AC1_TR = subset(Graph1,sex == "M" & acl00 =="AC1_TR" & age =="TOTAL")
time1 <- data_AC1_TR$TIME_SP
my_list_AC1_TR <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC1_TR <- mean(my_list_AC1_TR)

data_AC2 = subset(Graph1,sex == "M" & acl00 =="AC2" & age =="TOTAL")
time1 <- data_AC2$TIME_SP
my_list_AC2 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC2 <- mean(my_list_AC2)

data_AC3 = subset(Graph1,sex == "M" & acl00 =="AC3" & age =="TOTAL")
time1 <- data_AC3$TIME_SP
my_list_AC3 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC3 <- mean(my_list_AC3)

data_AC4_8 = subset(Graph1,sex == "M" & acl00 =="AC4-8" & age =="TOTAL")
time1 <- data_AC4_8$TIME_SP
my_list_AC4_8 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC4_8 <- mean(my_list_AC4_8)

myplot <- c(AC0, AC1_TR, AC2, AC3, AC4_8)
myplot <- gsub("Time difference of ","", myplot)
myplot <- gsub(" mins","", myplot)
myplot <- as.double(myplot)

#-------------------------------------------------------------------------------------------
#female activity time spent
data_AC0 = subset(Graph1,sex == "F" & acl00 =="AC0" & age =="TOTAL")
time1 <- data_AC0$TIME_SP
my_list_AC0 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC0 <- mean(my_list_AC0)

data_AC1_TR = subset(Graph1,sex == "F" & acl00 =="AC1_TR" & age =="TOTAL")
time1 <- data_AC1_TR$TIME_SP
my_list_AC1_TR <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC1_TR <- mean(my_list_AC1_TR)

data_AC2 = subset(Graph1,sex == "F" & acl00 =="AC2" & age =="TOTAL")
time1 <- data_AC2$TIME_SP
my_list_AC2 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC2 <- mean(my_list_AC2)

data_AC3 = subset(Graph1,sex == "F" & acl00 =="AC3" & age =="TOTAL")
time1 <- data_AC3$TIME_SP
my_list_AC3 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC3 <- mean(my_list_AC3)

data_AC4_8 = subset(Graph1,sex == "F" & acl00 =="AC4-8" & age =="TOTAL")
time1 <- data_AC4_8$TIME_SP
my_list_AC4_8 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC4_8 <- mean(my_list_AC4_8)

myplot2 <- c(AC0, AC1_TR, AC2, AC3, AC4_8)
myplot2 <- gsub("Time difference of ","", myplot2)
myplot2 <- gsub(" mins","", myplot2)
myplot2 <- as.double(myplot2)

#----------------------------------------------------------------------------------
#plot femals vs males time spent activities
par(mfrow = c(1:2))
g3 <- barplot(myplot,
              main = "Male activity time spent",
              xlab = "Activities",
              ylab = "Minutes",
              names.arg = c("Personal care", "Employment activities", "Study", "Family care", "Leisure time"),
              col = "lightblue", horiz=TRUE
              
              
)

g4 <- barplot(myplot2,
              main = "Female activity time spent",
              xlab = "Activities",
              ylab = "Minutes",
              names.arg = c("Personal care", "Employment activities", "Study", "Family care", "Leisure time"),
              col = "lightpink", horiz=TRUE
              
              
)

```


As we can see from the above graph and the one previously presented, in general women spend more time on activities related to family care, while males usually spend more time at employment related activities.

Since the biggest difference between males and females activities was regarding family/household related activities, we decided to make a graph about the secondary types of activities of this category.

Here we have a graph that shows the mean time spent by Male and Female counterparts on household related activities:


```{r maligraph3, echo=FALSE, fig.height=15,error=FALSE,warning=FALSE}
#---------------------------------------------------------------
#male activity time spent
data_AC31A = subset(Graph1,sex == "M" & acl00 =="AC31A" & age =="TOTAL")
time1 <- data_AC31A$TIME_SP
my_list_AC31A <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC31A <- mean(my_list_AC31A)

data_AC313 = subset(Graph1,sex == "M" & acl00 =="AC313" & age =="TOTAL")
time1 <- data_AC313$TIME_SP
my_list_AC313 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC313 <- mean(my_list_AC313)

data_AC331 = subset(Graph1,sex == "M" & acl00 =="AC331" & age =="TOTAL")
time1 <- data_AC2$TIME_SP
my_list_AC331 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC331 <- mean(my_list_AC331)

data_AC332 = subset(Graph1,sex == "M" & acl00 =="AC332" & age =="TOTAL")
time1 <- data_AC332$TIME_SP
my_list_AC332 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC332 <- mean(my_list_AC332)

data_AC34A = subset(Graph1,sex == "M" & acl00 =="AC34A" & age =="TOTAL")
time1 <- data_AC34A$TIME_SP
my_list_AC34A <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC34A <- mean(my_list_AC34A)

data_AC35 = subset(Graph1,sex == "M" & acl00 =="AC35" & age =="TOTAL")
time1 <- data_AC35$TIME_SP
my_list_AC35 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC35 <- mean(my_list_AC35)

myplot5 <- c(AC31A, AC313, AC331, AC332, AC34A, AC35)
myplot5 <- gsub("Time difference of ","", myplot5)
myplot5 <- gsub(" mins","", myplot5)
myplot5 <- as.double(myplot5)

#----------------------------------------------------------------------------------


#-------------------------------------------------------------------------------------------
#female activity time spent
data_AC31A = subset(Graph1,sex == "F" & acl00 =="AC31A" & age =="TOTAL")
time1 <- data_AC31A$TIME_SP
my_list_AC31A <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC31A <- mean(my_list_AC31A)

data_AC313 = subset(Graph1,sex == "F" & acl00 =="AC313" & age =="TOTAL")
time1 <- data_AC313$TIME_SP
my_list_AC313 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC313 <- mean(my_list_AC313)

data_AC331 = subset(Graph1,sex == "F" & acl00 =="AC331" & age =="TOTAL")
time1 <- data_AC2$TIME_SP
my_list_AC331 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC331 <- mean(my_list_AC331)

data_AC332 = subset(Graph1,sex == "F" & acl00 =="AC332" & age =="TOTAL")
time1 <- data_AC332$TIME_SP
my_list_AC332 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC332 <- mean(my_list_AC332)

data_AC34A = subset(Graph1,sex == "F" & acl00 =="AC34A" & age =="TOTAL")
time1 <- data_AC34A$TIME_SP
my_list_AC34A <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC34A <- mean(my_list_AC34A)

data_AC35 = subset(Graph1,sex == "F" & acl00 =="AC35" & age =="TOTAL")
time1 <- data_AC35$TIME_SP
my_list_AC35 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC35 <- mean(my_list_AC35)

myplot6 <- c(AC31A, AC313, AC331, AC332, AC34A, AC35)
myplot6 <- gsub("Time difference of ","", myplot6)
myplot6 <- gsub(" mins","", myplot6)
myplot6 <- as.double(myplot6)

#----------------------------------------------------------------------------------
#---------------------------------------------------
#plot females vs males time spent
par(mfrow = c(1:2))
g5 <- barplot(myplot5,
        main = "Male activity time spent",
        xlab = "Activities",
        ylab = "Minutes",
        names.arg = c("Food Management", "Dish washing", "Laundry", "Ironing", "Gardening", "Construction and repairs"),
        col = "lightblue", horiz = TRUE
        
        
)

g6 <- barplot(myplot6,
              main = "Female activity time spent",
              xlab = "Activities",
              ylab = "Minutes",
              names.arg = c("Food Management", "Dish washing", "Laundry", "Ironing", "Gardening", "Construction and repairs"),
              col = "lightpink", horiz = TRUE
              
              
)
```

As we can see from the graph above, in general women spend more time on Ironing, Dish washing and Food management, while males usually spend more time on Construction and repair related activities.

Here we have a graph that shows the mean time spent by Male and Female counterparts on leisure time related activities:

```{r maligraph4, echo=FALSE, fig.height=15,error=FALSE,warning=FALSE}
#male activity time spent
data_AC6A = subset(Graph1,sex == "M" & acl00 =="AC6A" & age =="TOTAL")
time1 <- data_AC6A$TIME_SP
my_list_AC6A <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC6A <- mean(my_list_AC6A)

data_AC733 = subset(Graph1,sex == "M" & acl00 =="AC733" & age =="TOTAL")
time1 <- data_AC733$TIME_SP
my_list_AC733 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC733 <- mean(my_list_AC733)

data_AC812 = subset(Graph1,sex == "M" & acl00 =="AC812" & age =="TOTAL")
time1 <- data_AC812$TIME_SP
my_list_AC812 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC812 <- mean(my_list_AC812)

data_AC82 = subset(Graph1,sex == "M" & acl00 =="AC82" & age =="TOTAL")
time1 <- data_AC82$TIME_SP
my_list_AC82 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC82 <- mean(my_list_AC82)

data_AC83 = subset(Graph1,sex == "M" & acl00 =="AC83" & age =="TOTAL")
time1 <- data_AC83$TIME_SP
my_list_AC83 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC83 <- mean(my_list_AC83)

data_AC7B = subset(Graph1,sex == "M" & acl00 =="AC7B" & age =="TOTAL")
time1 <- data_AC7B$TIME_SP
my_list_AC7B <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC7B <- mean(my_list_AC7B)

myplot7 <- c(AC6A, AC733, AC812, AC82, AC83, AC7B)
myplot7 <- gsub("Time difference of ","", myplot7)
myplot7 <- gsub(" mins","", myplot7)
myplot7 <- as.double(myplot7)
#female activity time spent
data_AC6A = subset(Graph1,sex == "F" & acl00 =="AC6A" & age =="TOTAL")
time1 <- data_AC6A$TIME_SP
my_list_AC6A <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC6A <- mean(my_list_AC6A)

data_AC733 = subset(Graph1,sex == "F" & acl00 =="AC733" & age =="TOTAL")
time1 <- data_AC733$TIME_SP
my_list_AC733 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC733 <- mean(my_list_AC733)

data_AC812 = subset(Graph1,sex == "F" & acl00 =="AC812" & age =="TOTAL")
time1 <- data_AC812$TIME_SP
my_list_AC812 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC812 <- mean(my_list_AC812)

data_AC82 = subset(Graph1,sex == "F" & acl00 =="AC82" & age =="TOTAL")
time1 <- data_AC82$TIME_SP
my_list_AC82 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC82 <- mean(my_list_AC82)

data_AC83 = subset(Graph1,sex == "F" & acl00 =="AC83" & age =="TOTAL")
time1 <- data_AC83$TIME_SP
my_list_AC83 <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC83 <- mean(my_list_AC83)

data_AC7B = subset(Graph1,sex == "F" & acl00 =="AC7B" & age =="TOTAL")
time1 <- data_AC7B$TIME_SP
my_list_AC7B <- difftime(as.POSIXct(time1, format = '%H:%M'), as.POSIXct('00:00', format = '%H:%M'), units = 'min')
AC7B <- mean(my_list_AC7B)

myplot8 <- c(AC6A, AC733, AC812, AC82, AC83, AC7B)
myplot8 <- gsub("Time difference of ","", myplot8)
myplot8 <- gsub(" mins","", myplot8)
myplot8 <- as.double(myplot8)

#-----------------------------------------------------------------------------------------------------------------
# plot time spent females vs males
par(mfrow = c(1:2))
g7 <- barplot(myplot7,
              main = "Male activity time spent",
              xlab = "Activities",
              ylab = "Minutes",
              names.arg = c("Sports", "Computer games", "Reading books", "TV and Video", "Music", "Hobbies"),
              col = "lightblue", horiz = TRUE
              
              
)
g8 <- barplot(myplot8,
        main = "Female activity time spent",
        xlab = "Activities",
        ylab = "Minutes",
        names.arg = c("Sports", "Computer games", "Reading books", "TV and Video", "Music", "Hobbies"),
        col = "lightpink", horiz = TRUE
        
        
)
```

As we can see from the above graph, in general women spend more time reading books, while males usually spend more time on computer games and sports .

